Kindly Bent to Free Us

### 2019-07-09 kindly/soa.tex
SOA.

Linear Haskell~\citep{DBLP:journals/pacmpl/BernardyBNJS18} aims
to retrofit Haskell with linear types.
Linear Haskell introduces a new notion
of ``linear arrows'', noted $\multimap$ similar to linear logic,
which are functions who use their argument exactly once.
By contrast, in \lang linearity is only decided
by the kinds, and $\tarr{\klin}$ simply denotes \emph{single-use functions}.
This design, while easy to retrofit on top of an existing compiler
such as GHC, has several disadvantages,
as highlighted by the long discussion in the GHC proposal\footnote{
  See the in-depth discussion starting here: \url{https://github.com/ghc-proposals/ghc-proposals/pull/111\#issuecomment-403349707}}.
Most relevant to \lang:
\begin{itemize}
\item Their system does not admit subtyping for arrows and leverages
  $\eta$-expansion to allow passing unrestricted functions in linear
  contexts. This is acceptable in a non-strict language such as
  Haskell but can not be used in a strict setting.
\item
  While the GHC proposal implements \emph{some} type inference,
  their algorithm is not formalized and not proved complete.
\item
  Linear Haskell promotes a ``bracket pattern'', similar to CPS-style such as 
  \lstinline/withFile : path -> (file ->. file) -> unit/.
  \lang follows System F\degree, Quill and Alms and
  adopts a more direct style, where types themselves are
  described as affine/linear.
  We believe this leads to a more natural programming style
  and makes reasoning about linearity in the context of abstraction much easier.
% \item
%   Due to the previous remark, ensuring that a given object is never
%   aliased is fairly difficult in Linear Haskell, as one would need to ensure
%   that no non-linear function can manipulate it. On the other
\end{itemize}

>>>>

Linear Haskell~\citep{DBLP:journals/pacmpl/BernardyBNJS18} aims
to retrofit Haskell with linear types.
Linear Haskell introduces a new notion
of ``linear arrows'', noted $\multimap$ similar to linear logic,
which are functions who use their argument exactly once.
By contrast, in \lang linearity is only decided
by the kinds, and $\tarr{\klin}$ simply denotes \emph{single-use functions}.
This design is easy to retrofit on top of an existing compiler
such as GHC, but has several disadvantages\footnote{
  See the in-depth discussion attached to the GHC proposal starting here: \url{https://github.com/ghc-proposals/ghc-proposals/pull/111\#issuecomment-403349707}}.
Most relevant to \lang:
\begin{itemize}
\item Their system does not admit subtyping for arrows and leverages
  $\eta$-expansion to allow passing unrestricted functions in linear
  contexts. This is acceptable in a non-strict language such as
  Haskell but can not be used in a strict setting.
\item
  While the GHC proposal implements some type inference,
  their algorithm is not formalized and not proved complete.
\item
  Linear Haskell promotes a CPS-style such as 
  \lstinline/withFile : path -> (file ->. file) ->. unit/
  to ensure that files are only used linearly.
  \lang follows System F\degree, Quill and Alms and
  adopts a direct style, where types themselves are
  described as affine/linear.
  We believe this leads to a more natural programming style
  and ease modular reasoning about linearity.
  In particular, using abstraction through modules,
  programmers only need consider the module
  implementation to ensure that some linear
  resource is properly handled.
\item
  Linear Haskell introduces a notion of linear monad to express
  imperative code conveniently. Again, this solution is suitable in Haskell,
  but less appropriate in other contexts. Our borrow system allow
  to provide a more free-form programming style for imperative code.
% \item
%   Due to the previous remark, ensuring that a given object is never
%   aliased is fairly difficult in Linear Haskell, as one would need to ensure
%   that no non-linear function can manipulate it. On the other
\end{itemize}
